Rbans-output

### 2021-07-24 reports/manuscript.Rmd
Updates manuscript with first review

```{r}
df_demographics_good %>%
  select(age, ilness_duration_years) %>%
  pivot_longer(everything()) %>%
  group_by(name) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  knitr::kable(digits = 3)

@@

```{r}
df_demographics_good %>%
  mutate(diagnosis = sapply(diagnosis, function(x){
    if(grepl("F3.*", x)) return("affective")
    if(grepl("F2.*", x)) return("schizophrenia")
  })) %>%
  count(diagnosis, first_training) %>%
  pivot_wider(names_from = diagnosis, values_from = n) %>%
  knitr::kable()

@@

df_rbans_long %>%
  filter(grepl("iq_", subtest), session == 1) %>%
  group_by(subtest, diagnosis) %>%
  summarise(m = mean_and_sd_report(score, type = "parentheses")) %>%
  pivot_wider(names_from = diagnosis, values_from = m) %>%
  right_join(df_diag_difference, by = "subtest") %>%
  knitr::kable(digits = 3)

df_rbans %>%
  group_by()

@@

```{r}
df_results_remediation %>%
  select(rbansid, session, max_difficulty) %>%
  distinct() %>%
  ggplot(aes(session, max_difficulty)) + 
    geom_jitter(height = 0, width = 0.15) +
    geom_smooth(method = "lm", fill = CDZ_COLORS[4], color = CDZ_COLORS[3]) +
    scale_x_continuous(breaks = seq(2,12,1)) + 
    theme_apa() +
    labs(y = "Maximum difficulty achieved", x = "Session")
ggsave("figures/fig1-session-difficulty.png",  width = 8, height = 6)
```

@@

```{r}
df_rbans_long %>%
  filter(grepl("iq_", subtest)) %>%
  mutate(session = paste0("session ", session - 1),
         fac_subtest = factor(subtest, levels = c("iq_celek", "iq_kratkodobapamet",
                                                  "iq_dlouhodobapamet", "iq_rec",
                                                  "iq_pozornost", "iq_vizuoprostorschopnosti"))) %>%
  ggplot(aes(fac_subtest, score, fill = trenink)) +
    geom_boxplot() +
    facet_wrap(~session,  strip.position = "bottom") +
    scale_x_discrete(labels = c("iq_celek" = "Total", "iq_kratkodobapamet" = "Immediate memory", 
      "iq_vizuoprostorschopnosti" = "Visuospatial/constructional", "iq_rec" = "Language", 
      "iq_pozornost" = "Attention", "iq_dlouhodobapamet" = "Delayed memory")) +
    scale_fill_manual(labels = c("baseline", "TP" = "Standard treatment",
                                   "VR" = "VEs program"),
                        values = CDZ_COLORS) +
    theme_apa() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = c(0.5,1), legend.direction = "horizontal") +
    labs(x = "", y = "RBANS index score",
         title = "Index scores of RBANS cognitive domains",
         fill = "")

@@

@@

@@

dims <- c("dim_celek", "dim_kratkodobapamet", "dim_dlouhodobapamet",
          "dim_pozornost", "dim_rec", "dim_vizuoprostor")
res <- sapply(dims, function(x){report_mixed_effect(temp_mixed_func(x), 3)})
knitr::kable(as.data.frame(res),
             caption = "Diagnosis as a predictor of RBANS performance. 
             Coefficients are for schizophrenia in contrast to MDD (negative = schizophrenia does worse)")
```
```{r}
df_rbans_long %>%
  filter(grepl("iq_", subtest)) %>%
  mutate(session = paste0("session ", session - 1),
         fac_subtest = factor(subtest, levels = c("iq_celek", "iq_kratkodobapamet",
                                                  "iq_dlouhodobapamet", "iq_rec",
                                                  "iq_pozornost", "iq_vizuoprostorschopnosti"))) %>%
  ggplot(aes(subtest, score, fill = trenink)) +
    scale_x_discrete(labels = c("iq_celek" = "Total", "iq_kratkodobapamet" = "Immediate memory", 
    "iq_vizuoprostorschopnosti" = "Visuospatial/constructional", "iq_rec" = "Language", 
    "iq_pozornost" = "Attention", "iq_dlouhodobapamet" = "Delayed memory")) +
    geom_boxplot() + 
    theme_apa() +
    scale_fill_manual(labels = c("baseline", "TP" = "Standard treatment",
                                   "VR" = "VEs program"),
                        values = CDZ_COLORS) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = c(0.5,1), legend.direction = "horizontal") +
      labs(x = "", y = "RBANS index score",
         title = "Index scores of RBANS cognitive domains after each intervention",
         fill = "")

ggsave("figures/fig3-rbans-iq.png",  width = 8, height = 6)
```

@@

>>>>

```{r demographics tables}
df_demographics_good %>%
  select(age, ilness_duration_years) %>%
  pivot_longer(everything()) %>%
  group_by(name) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  knitr::kable(digits = 3)

@@

```{r schizophrenia vs MDD demographics}
df_demographics_good %>%
  mutate(diagnosis = sapply(diagnosis, function(x){
    if(grepl("F3.*", x)) return("affective")
    if(grepl("F2.*", x)) return("schizophrenia")
  })) %>%
  count(diagnosis, first_training) %>%
  pivot_wider(names_from = diagnosis, values_from = n) %>%
  knitr::kable()

@@

df_rbans_long %>%
  filter(grepl("iq_", subtest), session == 1) %>%
  group_by(subtest, diagnosis) %>%
  summarise(m = mean_and_sd_report(score, type = "parentheses")) %>%
  pivot_wider(names_from = diagnosis, values_from = m) %>%
  right_join(df_diag_difference, by = "subtest") %>%
  knitr::kable(digits = 3)
```

@@

```{r}

df_results_remediation %>%
  select(rbansid, session, max_difficulty) %>%
  distinct() %>%
  ggplot(aes(session, max_difficulty)) + 
    geom_jitter(height = 0, width = 0.15) +
    geom_smooth(method = "lm", fill = CDZ_COLORS[4], color = CDZ_COLORS[3]) +
    scale_x_continuous(breaks = seq(2,12,1)) + 
    theme_apa() +
    labs(y = "Maximum difficulty achieved", x = "Session")
ggsave("figures/fig1-session-difficulty.png",  width = 8, height = 6)

df_results_remediation %>%
  select(rbansid, session, max_difficulty) %>%
  distinct() %>%
  ggplot(aes(session, max_difficulty)) + 
    geom_jitter(height = 0, width = 0.15) +
    geom_smooth(method = "lm", fill = "grey80", color = "grey20") +
    scale_x_continuous(breaks = seq(2,12,1)) + 
    theme_apa() +
    labs(y = "Maximum difficulty achieved", x = "Session")
ggsave("figures/fig1-session-difficulty-bw.png",  width = 8, height = 6)

@@

```{r}
fig2_rbans_iq <- df_rbans_long %>%
  filter(grepl("iq_", subtest)) %>%
  mutate(session = paste0("session ", session - 1),
         fac_subtest = factor(subtest, levels = c("iq_celek", "iq_kratkodobapamet",
                                                  "iq_dlouhodobapamet", "iq_rec",
                                                  "iq_pozornost", "iq_vizuoprostorschopnosti"))) %>%
  ggplot(aes(fac_subtest, score, fill = trenink)) +
    geom_boxplot() +
    facet_wrap(~session,  strip.position = "bottom") +
    scale_x_discrete(labels = c("iq_celek" = "Total", "iq_kratkodobapamet" = "Immediate memory", 
      "iq_vizuoprostorschopnosti" = "Visuospatial/constructional", "iq_rec" = "Language", 
      "iq_pozornost" = "Attention", "iq_dlouhodobapamet" = "Delayed memory")) +

    theme_apa() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = c(0.5,1), legend.direction = "horizontal") +
    labs(x = "", y = "RBANS index score",
         title = "Index scores of RBANS cognitive domains",
         fill = "")

fig2_rbans_iq + 
  scale_fill_manual(labels = c("baseline", "TP" = "Standard treatment",
                               "VR" = "VEs program"),
                        values = CDZ_COLORS)

@@

fig2_rbans_iq + 
  scale_fill_manual(labels = c("baseline", "TP" = "Standard treatment",
                               "VR" = "VEs program"),
                        values = colorRampPalette(c("white","grey40","grey80"))(3))

ggsave("figures/fig2-rbans-iq-bw.png", width = 8, height = 6)

@@

lmer_celek <- temp_mixed_func("dim_celek")
lmer_kratkodobapamet <- temp_mixed_func("dim_kratkodobapamet")
lmer_dlouhodobapamet <- temp_mixed_func("dim_dlouhodobapamet")
lmer_pozornost <- temp_mixed_func("dim_pozornost")
lmer_rec <- summary(temp_mixed_func("dim_rec"))
lmer_vizuoprostor <- temp_mixed_func("dim_vizuoprostor")

@@

dims <- c("dim_celek", "dim_kratkodobapamet", "dim_dlouhodobapamet",
          "dim_pozornost", "dim_rec", "dim_vizuoprostor")
res <- sapply(dims, function(x){report_mixed_effect(temp_mixed_func(x), 3)})
knitr::kable(as.data.frame(res),
             caption = "Diagnosis as a predictor of RBANS performance. 
             Coefficients are for schizophrenia in contrast to MDD (negative = schizophrenia does worse)")
```

```{r}
fig3_rbans_iq <- df_rbans_long %>%
  filter(grepl("iq_", subtest)) %>%
  mutate(session = paste0("session ", session - 1),
         fac_subtest = factor(subtest, levels = c("iq_celek", "iq_kratkodobapamet",
                                                  "iq_dlouhodobapamet", "iq_rec",
                                                  "iq_pozornost", "iq_vizuoprostorschopnosti"))) %>%
  ggplot(aes(subtest, score, fill = trenink)) +
    scale_x_discrete(labels = c("iq_celek" = "Total", "iq_kratkodobapamet" = "Immediate memory", 
    "iq_vizuoprostorschopnosti" = "Visuospatial/constructional", "iq_rec" = "Language", 
    "iq_pozornost" = "Attention", "iq_dlouhodobapamet" = "Delayed memory")) +
    geom_boxplot() + 
    theme_apa() +

    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = c(0.5,1), legend.direction = "horizontal") +
      labs(x = "", y = "RBANS index score",
         title = "Index scores of RBANS cognitive domains after each intervention",
         fill = "")

fig3_rbans_iq + scale_fill_manual(labels = c("baseline", "TP" = "Standard treatment",
                                            "VR" = "VEs program"), values = CDZ_COLORS)
ggsave("figures/fig3-rbans-iq.png", width = 8, height = 6)

fig3_rbans_iq + scale_fill_manual(labels = c("baseline", "TP" = "Standard treatment",
                                            "VR" = "VEs program"),
                                  values = colorRampPalette(c("white","grey40","grey80"))(3))
ggsave("figures/fig3-rbans-iq-bw.png",width = 8, height = 6)
```

@@

## Review 1.
The reviewer asked for separation of Schizophrenia and MDD patients, given their different cognitive decline/performance

```{r diagnosis as a predictor sessions 2-3}
temp_mixed_func <- function(varname){
  out <- df_rbans_long %>%
    filter(subtest == varname, session > 1) %>%
    lmer(z_score ~ diagnosis  + factor(session) + (1|rbansid), data = .)
  return(out)
}

lmer_celek <- temp_mixed_func("dim_celek")
lmer_kratkodobapamet <- temp_mixed_func("dim_kratkodobapamet")
lmer_dlouhodobapamet <- temp_mixed_func("dim_dlouhodobapamet")
lmer_pozornost <- temp_mixed_func("dim_pozornost")
lmer_rec <- summary(temp_mixed_func("dim_rec"))
lmer_vizuoprostor <- temp_mixed_func("dim_vizuoprostor")

dims <- c("dim_celek", "dim_kratkodobapamet", "dim_dlouhodobapamet",
          "dim_pozornost", "dim_rec", "dim_vizuoprostor")
res <- sapply(dims, function(x){report_mixed_effect(temp_mixed_func(x), 3)})
knitr::kable(as.data.frame(res),
             caption = "Diagnosis as a predictor of RBANS performance in sessions 2 and 3. 
             Coefficients are for schizophrenia in contrast to MDD (negative = schizophrenia does worse)")
```

```{r diagnosis as a predictor all sessions}
temp_mixed_func <- function(varname){
  out <- df_rbans_long %>%
    filter(subtest == varname) %>%
    lmer(z_score ~ diagnosis + session + (1|rbansid), data = .)
  return(out)
}

lmer_celek <- temp_mixed_func("dim_celek")
lmer_kratkodobapamet <- temp_mixed_func("dim_kratkodobapamet")
lmer_dlouhodobapamet <- temp_mixed_func("dim_dlouhodobapamet")
lmer_pozornost <- temp_mixed_func("dim_pozornost")
lmer_rec <- summary(temp_mixed_func("dim_rec"))
lmer_vizuoprostor <- temp_mixed_func("dim_vizuoprostor")

dims <- c("dim_celek", "dim_kratkodobapamet", "dim_dlouhodobapamet",
          "dim_pozornost", "dim_rec", "dim_vizuoprostor")
res <- sapply(dims, function(x){report_mixed_effect(temp_mixed_func(x), "diagnosisschizophrenia")})
knitr::kable(as.data.frame(res),
             caption = "Diagnosis as a predictor of RBANS performance while controlling for the effect of participant and session.
             Coefficients are for schizophrenia in contrast to MDD (negative = schizophrenia does worse)")

res <- sapply(dims, function(x){report_mixed_effect(temp_mixed_func(x), "session")})
knitr::kable(as.data.frame(res),
             caption = "Session as a predictor of RBANS performance while controlling for the effect of participant and diagnosis.
             negative = patients do worse and worse")
```

```{r}
df_rbans_long %>%
  filter(subtest == "dim_celek") %>%
  ggplot(aes(session, z_score)) +
  geom_point(position=position_jitter(width=0.2)) +
  geom_smooth(method="lm") +
  scale_x_continuous(breaks = 1:3) +
  facet_wrap(~diagnosis)
```

```{r only MDD patient}
temp_mixed_func <- function(varname, diagnosis){
  out <- df_rbans_long %>%
    filter(subtest == varname, session > 1, diagnosis == "affective disorders") %>%
    lmer(z_score ~ factor(session) + (1|rbansid), data = .)
  return(out)
}

lmer_celek <- temp_mixed_func("dim_celek")
lmer_kratkodobapamet <- temp_mixed_func("dim_kratkodobapamet")
lmer_dlouhodobapamet <- temp_mixed_func("dim_dlouhodobapamet")
lmer_pozornost <- temp_mixed_func("dim_pozornost")
lmer_rec <- summary(temp_mixed_func("dim_rec"))
lmer_vizuoprostor <- temp_mixed_func("dim_vizuoprostor")

dims <- c("dim_celek", "dim_kratkodobapamet", "dim_dlouhodobapamet",
          "dim_pozornost", "dim_rec", "dim_vizuoprostor")
res <- sapply(dims, function(x){report_mixed_effect(temp_mixed_func(x), "factor(session)3")})
knitr::kable(as.data.frame(res),
             caption = "Session as a predictor of RBANS performance for patients in MDD. 
             Coefficients are for third session in contrast to the second (negative = patients do worse in 3rd session)")
```

```{r only SZ patient}
temp_mixed_func <- function(varname){
  out <- df_rbans_long %>%
    filter(subtest == varname, session > 1, diagnosis == "schizophrenia") %>%
    lmer(z_score ~ factor(session) + (1|rbansid), data = .)
  return(out)
}

lmer_celek <- temp_mixed_func("dim_celek")
lmer_kratkodobapamet <- temp_mixed_func("dim_kratkodobapamet")
lmer_dlouhodobapamet <- temp_mixed_func("dim_dlouhodobapamet")
lmer_pozornost <- temp_mixed_func("dim_pozornost")
lmer_rec <- summary(temp_mixed_func("dim_rec"))
lmer_vizuoprostor <- temp_mixed_func("dim_vizuoprostor")

dims <- c("dim_celek", "dim_kratkodobapamet", "dim_dlouhodobapamet",
          "dim_pozornost", "dim_rec", "dim_vizuoprostor")
res <- sapply(dims, function(x){report_mixed_effect(temp_mixed_func(x), "factor(session)3")})
knitr::kable(as.data.frame(res),
             caption = "Session as a predictor of RBANS performance for patients with schizophrenia. 
             Coefficients are for third session in contrast to the second (negative = patients do worse in 3rd session)")
```

```{r only SZ patient linear session}
temp_mixed_func <- function(varname){
  out <- df_rbans_long %>%
    filter(subtest == varname, diagnosis == "schizophrenia") %>%
    lmer(z_score ~ session + (1|rbansid), data = .)
  return(out)
}

lmer_celek <- temp_mixed_func("dim_celek")
lmer_kratkodobapamet <- temp_mixed_func("dim_kratkodobapamet")
lmer_dlouhodobapamet <- temp_mixed_func("dim_dlouhodobapamet")
lmer_pozornost <- temp_mixed_func("dim_pozornost")
lmer_rec <- summary(temp_mixed_func("dim_rec"))
lmer_vizuoprostor <- temp_mixed_func("dim_vizuoprostor")

dims <- c("dim_celek", "dim_kratkodobapamet", "dim_dlouhodobapamet",
          "dim_pozornost", "dim_rec", "dim_vizuoprostor")
res <- sapply(dims, function(x){report_mixed_effect(temp_mixed_func(x), 2)})
knitr::kable(as.data.frame(res),
             caption = "Session as a predictor of RBANS performance for patients with schizophrenia. 
             Coefficients are for session as a linear predictor (1,2,3) negative = patients do worse and worse")
```

```{r only MDD patient linear session}
temp_mixed_func <- function(varname){
  out <- df_rbans_long %>%
    filter(subtest == varname, diagnosis == "affective disorders") %>%
    lmer(z_score ~ session + (1|rbansid), data = .)
  return(out)
}

lmer_celek <- temp_mixed_func("dim_celek")
lmer_kratkodobapamet <- temp_mixed_func("dim_kratkodobapamet")
lmer_dlouhodobapamet <- temp_mixed_func("dim_dlouhodobapamet")
lmer_pozornost <- temp_mixed_func("dim_pozornost")
lmer_rec <- summary(temp_mixed_func("dim_rec"))
lmer_vizuoprostor <- temp_mixed_func("dim_vizuoprostor")

dims <- c("dim_celek", "dim_kratkodobapamet", "dim_dlouhodobapamet",
          "dim_pozornost", "dim_rec", "dim_vizuoprostor")
res <- sapply(dims, function(x){report_mixed_effect(temp_mixed_func(x), 2)})
knitr::kable(as.data.frame(res),
             caption = "Session as a predictor of RBANS performance for patients with affective disorders. 
             Coefficients are for session as a linear predictor (1,2,3) negative = patients do worse and worse")
```

```{r temp}
df_rbans_long %>%
  filter(diagnosis =="affective disorders", subtest=="dim_celek")

df_rbans %>%
  filter(diagnosis=="affective disorders", !is.na(testpameti)) %>%
  count(rbansid)
```

### 2021-07-24 reports/rbans-output.Rmd
Updates manuscript with first review

df_rbans_long <- df_rbans %>%
    pivot_longer(-c(rbansid, trenink, diagnosis, first_training, session),
                 names_to = "subtest", values_to = "score")

df_rbans_long <- df_rbans_long %>%
  group_by(trenink, first_training, subtest) %>%
  summarise(group_avg = mean(score, na.rm = TRUE)) %>%
  right_join(df_rbans_long, by = c("trenink", "first_training", "subtest"))
```
```{r}
df_ids %>%
  count(is_ok, rbans_ok, logy_ok) %>%
  kable(caption = "Overview of the state of affairs")

@@

>>>>

good_participants <- df_rbans %>%
  group_by(rbansid) %>%
  filter(!is.na(testpameti)) %>%
  summarise(n = n()) %>%
  filter(n >= 2) %>%
  pull(rbansid)

df_rbans <- df_rbans %>%
  filter(rbansid %in% good_participants)

df_rbans_long <- df_rbans %>%
    pivot_longer(-c(rbansid, trenink, diagnosis, first_training, session),
                 names_to = "subtest", values_to = "score") %>%
  group_by(session, subtest) %>%
  mutate(z_score = scale(score)[,1])

df_rbans_long <- df_rbans_long %>%
  group_by(trenink, first_training, subtest) %>%
  summarise(group_avg = mean(score, na.rm = TRUE),
            group_z_avg = mean(z_score, na.rm = TRUE),
            .groups="drop") %>%
  right_join(df_rbans_long, by = c("trenink", "first_training", "subtest"))
```

```{r}
df_ids %>%
  count(is_ok, rbans_ok, logy_ok) %>%
  kable(caption = "Overview of the state of affairs")

@@

```{r}
df_rbans_long %>%
  filter(!grepl("(iq_|dim_)", subtest)) %>%
  ggplot(aes(subtest, score, fill = factor(session))) + geom_boxplot() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = c(0.5,1), legend.direction = "horizontal") +
  facet_wrap(~trenink)
```

```{r}
df_rbans_long %>%
  filter(grepl("(iq_)", subtest)) %>%
  ggplot(aes(subtest, score, fill = factor(session))) + geom_boxplot() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = c(0.5,1), legend.direction = "horizontal")
```
